---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# traudem

<!-- badges: start -->
<!-- badges: end -->

The goal of traudem is to wrap the CLI TauDEM from R.

## Installation

Refer to `vignette("traudem-installation", package = "traudem")`.
In particular after installing the TauDEM CLI and dependencies as well as the R package, please run `traudem::taudem_sitrep()`.

```{r, echo=FALSE}
Sys.setenv(TAUDEM_PATH="/usr/local/taudem")
```

```{r}
traudem::taudem_sitrep()
```

You can install the development version of traudem from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cynkra/traudem")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(traudem)
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "MED_01_01.tif", package = "traudem"),
  file.path(test_dir, "MED_01_01.tif")
)
output <- taudem_pitremove(file.path(test_dir, "MED_01_01.tif"))
output
```
