---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# traudem

<!-- badges: start -->
<!-- badges: end -->

The goal of traudem is to wrap the [TauDEM CLI](https://hydrology.usu.edu/taudem/taudem5/index.html) from R.

## Installation

Refer to [`vignette("taudem-installation", package = "traudem")`](https://lucarraro.github.io/traudem/articles/taudem-installation.html).
In particular after installing the TauDEM CLI and dependencies as well as the R package, please run `traudem::taudem_sitrep()`.

```{r, echo=FALSE}
Sys.setenv(TAUDEM_PATH="/usr/local/taudem")
```

```{r}
traudem::taudem_sitrep()
```

You can install the development version of traudem from R-universe:

```r
# Enable repository from lucarraro
options(
  repos = c(
    lucarraro = 'https://lucarraro.r-universe.dev',
    CRAN = 'https://cloud.r-project.org'
  )
)
# Download and install traudem in R
install.packages('traudem')
```


Or from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lucarraro/traudem")
```

## Example

This is how you would use traudem to launch [TauDEM's PitRemove](https://hydrology.usu.edu/taudem/taudem5/help53/PitRemove.html):

```{r example}
library(traudem)
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"))
output
```

We ran the example above in a temporary directory that `withr` automatically deletes.
If you want to automatically get rid of some of the intermediary files created by TauDEM in one of your pipelines, you might be interested in `withr::local_tempfile()`.

If you wanted to run this same code without seeing the messages from TauDEM you can either use the `quiet` argument:

```{r quiet}
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"), quiet = TRUE)
output
```

Or set the `traudem.quiet` option (`options(traudem.quiet = TRUE)` or for just the session, `withr::local_options(traudem.quiet = TRUE)`):

```{r quietoption}
withr::local_options(traudem.quiet = TRUE)
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"))
output
```

## Can traudem run all TauDEM methods?

Yes!
Some TauDEM methods have dedicated wrappers like `taudem_pitremove()`, with argument names that we strived to make informative.
For other methods, you can use `taudem_exec()` and provide the arguments as TauDEM CLI would expect them.
You could make a PR to this repository to add more dedicated wrappers.
