---
title: "Install TauDEM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Install TauDEM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Installing TauDEM actually means installing 3 tools

* GDAL that TauDEM uses under the hood;
* MPI that R will call to launch TauDEM;
* TauDEM itself.

No matter your operating system, take notes of the steps you follow.
This way, if anything goes wrong, you will be able to open an informative bug report and get help more easily.

## Windows

You're in luck! 
The TauDEM Windows installer will install all three tools at once! 
Refer to <https://hydrology.usu.edu/taudem/taudem5/downloads.html>.

```{r, child="_sitrep.Rmd"}
```

## Linux and macOS

### GDAL

Do you need to install GDAL? In a terminal try `gdalinfo --version`, if it fails `gdal-config --version`. If it works and return something you do not need to install GDAL unless you want a newer version.
You might have installed GDAL in the past for, say, using the sf R package.

If you need to, install GDAL using the official [GDAL docs](https://gdal.org/download.html) (or refer to [TauDEM README](https://github.com/dtarb/TauDEM/) for further suggestions). 

> On macOS this might be as easy as brew install gdal.

### MPI

First try running `mpiexec --version` in a terminal.
If it works and return something you do not need to install MPI unless you want a newer version.

If you need to, install MPI using the official [MPI docs](https://www.open-mpi.org/software/ompi/v4.1/) (or refer to [TauDEM README](https://github.com/dtarb/TauDEM/) for further suggestions). 

> On macOS this might be as easy as brew install mpich.

### TauDEM

First you need to obtain TauDEM source for GitHub <https://github.com/dtarb/TauDEM/>.
If you are at ease with this, you can clone the repository.
If not, find the green button that allows you to clone, open (with GitHub CLI) or download, and download the ZIP version.

We recommend using CMake because that is what we did but [TauDEM README](https://github.com/dtarb/TauDEM/) for further suggestions) also has a method using Make.
Therefore,

* install CMake or Make (check it's not already installed);
* follow [TauDEM README](https://github.com/dtarb/TauDEM/) for further suggestions) for either CMake or Make depending on what you decided to use.

> If the installation shows errors, try using <https://github.com/maelle/TauDEM/tree/more-mpi-deprecations> instead.

```{r, child="_sitrep.Rmd"}
```
