## Simple example

This is how you would use traudem to launch [TauDEM's PitRemove](https://hydrology.usu.edu/taudem/taudem5/help53/PitRemove.html):

```{r example}
library(traudem)
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"))
output
```

We ran the example above in a temporary directory that `withr` automatically deletes.
If you want to automatically get rid of some of the intermediary files created by TauDEM in one of your pipelines, you might be interested in `withr::local_tempfile()`.

If you wanted to run this same code without seeing the messages from TauDEM, you can either use the `quiet` argument:

```{r quiet}
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"), quiet = TRUE)
output
```

Or set the `traudem.quiet` option (`options(traudem.quiet = TRUE)` or for just the session, `withr::local_options(traudem.quiet = TRUE)`):

```{r quietoption}
withr::local_options(traudem.quiet = TRUE)
test_dir <- withr::local_tempdir()
fs::file_copy(
  system.file("test-data", "DEM.tif", package = "traudem"),
  file.path(test_dir, "DEM.tif")
)
output <- taudem_pitremove(file.path(test_dir, "DEM.tif"))
output
```
