## Full example: download DEM data and extract a river network

`traudem` can be used in combination with R-package [elevatr](https://cran.r-project.org/package=elevatr) to extract a river network from elevation data retrieved from the web. 

The following example shows how to derive the contour of a catchment and drainage area values for all cells within that contour by using the [D8 flow direction algorithm](https://hydrology.usu.edu/taudem/taudem5/help53/D8FlowDirections.html).

The only required inputs are:

- coordinates of the catchment outlet;
- coordinates of the lower-left and top-right corners of a rectangular area containing the catchment of interest;
- EPSG code corresponding to the projection used to express coordinate data (see also https://epsg.org/home.html);
- zoom level at which the DEM data are extracted (see `elevatr` [documentation](https://cran.r-project.org/package=elevatr)).

```{r, fig.alt = " Plot of the catchment contour of the river Wigger (Switzerland) extracted from a DEM via traudem."}
library(elevatr)
library(sf)
library(terra)
library(traudem)
library(shapefiles)
library(withr)

taudem_sitrep() # verify that TauDEM is correctly installed 

test_dir <- local_tempdir() # temporary directory storing intermediary files created by TauDEM

# Input data for the river Wigger (central Switzerland)
EPSG <- 21781	# EPSG code corresponding to CH1903 (Swiss projected coordinate system) 
x_outlet <- 637478 # outlet x coordinate in CH1903 coordinates
y_outlet <- 237413 # outlet y coordinate in CH1903 coordinates
x_ll <- 620000 # x coordinate of the lower-left (SW) limit of the region
x_tr <- 660000 # x coordinate of the top-right (NE) limit of the region	
y_ll <- 200000 # y coordinate of the lower-left (SW) limit of the region	
y_tr <- 250000 # y coordinate of the top-right (NE) limit of the region
zoom <- 9 # corresponds to cellsize ~= 100 m at the latitude of interest

# use elevatr to download DEM 
loc.df <- data.frame(x=c(x_ll, x_tr), y=c(y_ll,y_tr))
r <- rast()
crs(r) <- paste0("epsg:", EPSG)
crs_str <- crs(r)

z <- get_elev_raster(locations = loc.df, prj = crs_str, z = zoom, verbose = F) 
z <- rast(z)
z <- crop(z, ext(x_ll, x_tr, y_ll, y_tr))
# reclassify: all pixels with elev=NA are set to 0. Then the pit remove algorithm will take care of them
z <- classify(z, matrix(c(NA, NA, 0), 1, 3)) 

writeRaster(z, filename = file.path(test_dir, "DEM.tif")) # write elevation raster to temporary directory

# apply TauDEM functions
# Remove pits
out_fel <- taudem_pitremove(file.path(test_dir, "DEM.tif"), quiet = TRUE)

# D8 flow directions
out_p <- taudem_d8flowdir(out_fel, quiet = TRUE)
out_p <- out_p$output_d8flowdir_grid # file path of flow direction file

# Contributing area
out_ad8 <- taudem_aread8(out_p, quiet = TRUE)

# Threshold
out_src <- taudem_threshold(out_ad8, quiet = TRUE)

# create shapefile with approximate outlet location
shp.point <- function(x, y, sname = "shape"){ # function to create point shapefile given coordinates 
  n <- length(x)
  dd <- data.frame(Id = 1:n, X = x, Y = y)
  ddTable <- data.frame(Id = c(1), Name = paste("Outlet", 1:n, sep = ""))
  ddShapefile <- convert.to.shapefile(dd, ddTable, "Id", 1)
  write.shapefile(ddShapefile, sname, arcgis = T)
  invisible(paste0(sname,".shp"))
}

out_shp <-  shp.point(x_outlet, y_outlet, file.path(test_dir, "ApproxOutlet"))

# Move outlet to stream
out_moved.shp <- taudem_moveoutletstostream(out_p, out_src, outlet_file = out_shp, 
                                            output_moved_outlets_file = file.path(test_dir, "Outlet.shp"), 
                                            quiet = TRUE)

# Contributing area upstream of outlet
out_ssa <- taudem_aread8(out_p, outlet_file = out_moved.shp, quiet = TRUE)

# Derive catchment contour
ssa <- rast(out_ssa)
# reclassify: pixels within the catchment have value 1, others 0
ssa_cont <- classify(ssa, matrix(c(NA, NA, 0, 1, Inf, 1), 2, 3, byrow = T)) 

# plot DEM of the region and catchment contours
plot(z, col = terrain.colors(1000), range = c(0,1200))
contour(ssa_cont, add = T, labels = "") 
title("DEM and catchment contour")
```


```
# plot map of drainage area within the catchment
plot(ssa, col = hcl.colors(1000, "viridis"))
title("Drainage area [no. cells]")
```


```{r, fig.alt = "Plot of log-scaled drainage area values for DEM pixels within the catchment. The use of a log-scale highlights the shape of the underlying river network."}
# total catchment area 
ssa_log <- ssa
values(ssa_log) <- log10(values(ssa))
plot(ssa_log, col = hcl.colors(1000, "viridis"))
title("Drainage area [log10(no. cells)]")

cellsize <- sqrt(prod(res(z))) # cellsize is geometric mean of downloaded raster resolution

max(values(ssa),na.rm=T)*cellsize^2 # total catchment area in m^2
```

The total area of the extracted catchment is `r round(max(values(ssa),na.rm=T)*cellsize^2/10^6)` km^2^.
